Concurrent Programming in Harmony

### 2021-04-16 book/paper.tex
Latest

>>>>

\begin{figure}
\begin{code}
\harmonysource{clock}
\end{code}
\caption{\harmonylink{code/clock.hny} Harmony program that finds page replacement anomalies.}
\label{fig:clock}
\end{figure}

If you are ready to learn about how locks are implemented in practice,
you can now skip the rest of this chapter.  But if you would like to
see a cool example of using the concepts introduced in this chapter,
hang on for a sequential Harmony program that finds anomalies in page
replacement algorithms.
In 1969, B\'{e}l\'{a}ady et al. published a paper~\cite{Belady69}
that showed that making a
cache larger does not necessarily lead to a higher hit ratio.  He showed
this for a scenario using a FIFO replacement policy when the cache is full. 
The program in \autoref{fig:clock} will find exactly the same scenario
if you define \texttt{FIFO} to be \textbf{True}.  Moreover, if you define
\texttt{FIFO} to be \textbf{False}, it will find a scenario for the
CLOCK replacement policy~\cite{clock}, often used in modern operating systems.

In this program, \texttt{CLOCK} maintains the state of a cache
(in practice, typically pages in memory).
The set \texttt{recent} maintains whether an access to the cache
for a particular reference was recent or not.
(It is not used if \texttt{FIFO} is \textbf{True}.)
The integer \texttt{misses} maintains the number of cache misses.
Method \texttt{ref}(\textit{ck}, $x$) is invoked when $x$ is
referenced and checked against the cache \textit{ck}.

The program declares two caches: one with 3 entries (\textit{clock3})
and one with 4 entries (\textit{clock4}).
The interesting part is in the last block of code.
It runs every sequence of references of up to 100 entries, using
references in the range 1 through 5.
Note that all the constants chosen in this program (3, 4, 5, 100)
are the result of some experimentation---you can try other ones.
To reduce the search space, the first four references are pinned
to 1, 2, 3, and 4.
Further reducing the search space, the program never repeats the
same reference twice in a row (using the local variable \texttt{last}).

The two things to note here is the use of the \textbf{choose}
expression and the \textbf{assert} statement.  Using \texttt{choose}
we are able to express searching through all possible strings of
references without a complicated nested iteration.
Using \texttt{assert} we are able to express the anomaly we are
looking for.

In case you want to check if you get the right results.  For \texttt{FIFO},
the program finds (in a few minutes) the same anomaly that
B\'{e}l\'{a}dy et al. found: 1 2 3 4 1 2 5 1 2 3 4 5.
For the \texttt{CLOCK} algorithm the program actually finds a shorter
reference string: 1 2 3 4 2 1 2 5 1 2.
